---
title: "plots"
author: "Chris Soria"
date: "2025-01-08"
output: html_document
---

```{r}
library(ggplot2)
library(dplyr)
library(purrr)
library(patchwork)
library(cowplot)
library(tidyverse)
```
# read in data
```{r setup, include=FALSE}
target_dir <- normalizePath("/Users/chrissoria/Documents/Research/us_international_ses/plots")
current_dir <- normalizePath(getwd())

if (current_dir == target_dir) {
  out_dir <- file.path("/Users/chrissoria/Documents/Research/us_international_ses/plots/")
  data_dir <- file.path("/Users/chrissoria/Documents/Research/us_international_ses/data/")
  plot_dir <- file.path("/Users/chrissoria/Documents/Research/us_international_ses//plots/")

} else {
  out_dir <- file.path("C:/Users/Ty/Desktop/ses_international/plots//")
  data_dir <- file.path("C:/Users/Ty/Desktop/ses_international/data//")
  plot_dir <- file.path("C:/Users/Ty/Desktop/ses_international/plots//")

}

print(getwd())
```

```{r}
us_df_2010 <- read_dta(paste0(data_dir, "US_2010_v100.dta"))
us_df_2020 <- read_dta(paste0(data_dir, "US_2020_v100.dta"))
```

## education in native and migrants
```{r}
countries <- c("Mexicans", "Puerto Ricans", "Dominicans", "Cubans", "US Mexican-born", "US Puerto-Rican-born", "US Dominican-born", "US Cuban-born")
education <- c("less_than_primary", "primary_completed", "secondary_completed", "university_completed", "unknown")

create_country_plot <- function(country, gender) {
  if(gender == "women") {
    prop_educated <- data.frame(
      levels = education,
      proportion = c(as.numeric(table1[5:8, country]), 1 - sum(as.numeric(table1[5:8, country]))) / 1
    )
    title_suffix <- "Women"
  } else {
    prop_educated <- data.frame(
      levels = education,
      proportion = c(as.numeric(table1[14:17, country]), 1 - sum(as.numeric(table1[14:17, country]))) / 1
    )
    title_suffix <- "Men"
  }
  
  ggplot(prop_educated, aes(x = "", y = proportion, fill = levels)) +
    geom_bar(stat = "identity", width = 0.5) +
    geom_text(aes(label = scales::percent(proportion, accuracy = 0.1)), 
              position = position_stack(vjust = 0.5), size = 3) +
    scale_y_continuous(labels = scales::percent) +
    scale_fill_manual(name = paste("Highest Education Achieved for", title_suffix),
      values = c("less_than_primary" = "skyblue", 
                 "primary_completed" = "lightgreen", 
                 "secondary_completed" = "red", 
                 "university_completed" = "orange",
                 "unknown" = "gray"),
      labels = c("Less than Primary", "Primary Completed", 
                 "Secondary Completed", "University Completed",
                 "Unknown")) +
    labs(title = paste(country, "-", title_suffix),
         y = element_blank(),
         x = NULL) +
    theme_cowplot() +
    theme(axis.text.x = element_blank(),
          axis.text.y = element_blank(),
          axis.ticks.x = element_blank(),
          axis.ticks.y = element_blank(),
          legend.position = "none",
          plot.title = element_text(hjust = 0.5),
          plot.subtitle = element_text(hjust = 0.5))
}

# Create plots for women
plots_women <- map(countries, ~create_country_plot(.x, "women"))

# Create plots for men
plots_men <- map(countries, ~create_country_plot(.x, "men"))

# Combine plots for women
combined_plot_women <- wrap_plots(plots_women, ncol = 4) +
  plot_annotation(
    title = "Education Levels by Country - Women",
    subtitle = "Based on table1",
    theme = theme(plot.title = element_text(hjust = 0.5),
                  plot.subtitle = element_text(hjust = 0.5))
  ) &
  theme(legend.position = "bottom")

combined_plot_women <- combined_plot_women + 
  plot_layout(guides = "collect") & 
  theme(legend.position = "bottom")

# Combine plots for men
combined_plot_men <- wrap_plots(plots_men, ncol = 4) +
  plot_annotation(
    title = "Education Levels by Country - Men",
    subtitle = "Based on table1",
    theme = theme(plot.title = element_text(hjust = 0.5),
                  plot.subtitle = element_text(hjust = 0.5))
  ) &
  theme(legend.position = "bottom")

combined_plot_men <- combined_plot_men + 
  plot_layout(guides = "collect") & 
  theme(legend.position = "bottom")

# Save plots
ggsave(filename = paste0(out_dir,"education_panels_women.jpg"), plot = combined_plot_women, width = 15, height = 10, dpi = 300)
ggsave(filename = paste0(out_dir,"education_panels_men.jpg"), plot = combined_plot_men, width = 15, height = 10, dpi = 300)

# Print plots
print(combined_plot_women)
print(combined_plot_men)
```
## percent married in native and migrant pop
```{r}
countries <- c("Mexicans", "Puerto Ricans", "Dominicans", "Cubans", "US Mexican-born", "US Puerto-Rican-born", "US Dominican-born", "US Cuban-born")
marital_status <- c("married", "not_married")

create_country_plot <- function(country, gender) {
  if(gender == "women") {
    prop_married <- data.frame(
      levels = marital_status,
      proportion = c(as.numeric(table1[4, country]), 1 - as.numeric(table1[4, country])) / 1
    )
    title_suffix <- "Women"
  } else {
    prop_married <- data.frame(
      levels = marital_status,
      proportion = c(as.numeric(table1[13, country]), 1 - as.numeric(table1[13, country])) / 1
    )
    title_suffix <- "Men"
  }
  
  ggplot(prop_married, aes(x = "", y = proportion, fill = levels)) +
    geom_bar(stat = "identity", width = 0.5) +
    geom_text(aes(label = scales::percent(proportion, accuracy = 0.1)), 
              position = position_stack(vjust = 0.5), size = 3) +
    scale_y_continuous(labels = scales::percent) +
    scale_fill_manual(name = paste("Marital Status -", title_suffix),
      values = c("married" = "lightblue", 
                 "not_married" = "pink"),
      labels = c("Married", "Not Married")) +
    labs(title = paste(country, "-", title_suffix),
         y = element_blank(),
         x = NULL) +
    theme_cowplot() +
    theme(axis.text.x = element_blank(),
          axis.text.y = element_blank(),
          axis.ticks.x = element_blank(),
          axis.ticks.y = element_blank(),
          legend.position = "none",
          plot.title = element_text(hjust = 0.5),
          plot.subtitle = element_text(hjust = 0.5))
}

# Create plots for women
plots_women <- map(countries, ~create_country_plot(.x, "women"))

# Create plots for men
plots_men <- map(countries, ~create_country_plot(.x, "men"))

# Combine plots for women
combined_plot_women <- wrap_plots(plots_women, ncol = 4) +
  plot_annotation(
    title = "Marital Status by Country - Women",
    subtitle = "Based on table1",
    theme = theme(plot.title = element_text(hjust = 0.5),
                  plot.subtitle = element_text(hjust = 0.5))
  ) &
  theme(legend.position = "bottom")

combined_plot_women <- combined_plot_women + 
  plot_layout(guides = "collect") & 
  theme(legend.position = "bottom")

# Combine plots for men
combined_plot_men <- wrap_plots(plots_men, ncol = 4) +
  plot_annotation(
    title = "Marital Status by Country - Men",
    subtitle = "Based on table1",
    theme = theme(plot.title = element_text(hjust = 0.5),
                  plot.subtitle = element_text(hjust = 0.5))
  ) &
  theme(legend.position = "bottom")

combined_plot_men <- combined_plot_men + 
  plot_layout(guides = "collect") & 
  theme(legend.position = "bottom")

# Save plots
ggsave(filename = paste0(out_dir,"marital_status_panels_women.jpg"), plot = combined_plot_women, width = 15, height = 10, dpi = 300)
ggsave(filename = paste0(out_dir,"marital_status_panels_men.jpg"), plot = combined_plot_men, width = 15, height = 10, dpi = 300)

# Print plots
print(combined_plot_women)
print(combined_plot_men)
```
## age groups in natives and migrants
```{r}
countries <- c("Mexicans", "Puerto Ricans", "Dominicans", "Cubans", "US Mexican-born", "US Puerto-Rican-born", "US Dominican-born", "US Cuban-born")
age_groups <- c("Age 60 - 69", "Age 70 - 79", "Age 80 - 89")

create_country_plot <- function(country, gender) {
  if(gender == "women") {
    prop_age <- data.frame(
      levels = age_groups,
      proportion = c(as.numeric(table1[1:3, country])) / 1
    )
    title_suffix <- "Women"
  } else {
    prop_age <- data.frame(
      levels = age_groups,
      proportion = c(as.numeric(table1[10:12, country])) / 1
    )
    title_suffix <- "Men"
  }
  
  ggplot(prop_age, aes(x = "", y = proportion, fill = levels)) +
    geom_bar(stat = "identity", width = 0.5) +
    geom_text(aes(label = scales::percent(proportion, accuracy = 0.1)), 
              position = position_stack(vjust = 0.5), size = 3) +
    scale_y_continuous(labels = scales::percent) +
    scale_fill_manual(name = paste("Age Groups -", title_suffix),
      values = c("Age 60 - 69" = "lightblue", 
                 "Age 70 - 79" = "orange",
                 "Age 80 - 89" = "lightgreen"),
      labels = age_groups) +
    labs(title = paste(country, "-", title_suffix),
         y = element_blank(),
         x = NULL) +
    theme_cowplot() +
    theme(axis.text.x = element_blank(),
          axis.text.y = element_blank(),
          axis.ticks.x = element_blank(),
          axis.ticks.y = element_blank(),
          legend.position = "none",
          plot.title = element_text(hjust = 0.5),
          plot.subtitle = element_text(hjust = 0.5))
}

# Create plots for women
plots_women <- map(countries, ~create_country_plot(.x, "women"))

# Create plots for men
plots_men <- map(countries, ~create_country_plot(.x, "men"))

# Combine plots for women
combined_plot_women <- wrap_plots(plots_women, ncol = 4) +
  plot_annotation(
    title = "Age Distribution by Country - Women",
    subtitle = "Based on table1",
    theme = theme(plot.title = element_text(hjust = 0.5),
                  plot.subtitle = element_text(hjust = 0.5))
  ) &
  theme(legend.position = "bottom")

combined_plot_women <- combined_plot_women + 
  plot_layout(guides = "collect") & 
  theme(legend.position = "bottom")

# Combine plots for men
combined_plot_men <- wrap_plots(plots_men, ncol = 4) +
  plot_annotation(
    title = "Age Distribution by Country - Men",
    subtitle = "Based on table1",
    theme = theme(plot.title = element_text(hjust = 0.5),
                  plot.subtitle = element_text(hjust = 0.5))
  ) &
  theme(legend.position = "bottom")

combined_plot_men <- combined_plot_men + 
  plot_layout(guides = "collect") & 
  theme(legend.position = "bottom")

# Save plots
ggsave(filename = paste0(out_dir,"age_distribution_panels_women.jpg"), plot = combined_plot_women, width = 15, height = 10, dpi = 300)
ggsave(filename = paste0(out_dir,"age_distribution_panels_men.jpg"), plot = combined_plot_men, width = 15, height = 10, dpi = 300)

# Print plots
print(combined_plot_women)
print(combined_plot_men)

```

## english speakers in migrant population
```{r}
countries <- c("mexico", "puerto rico", "dominican republic", "cuba")

result_list <- list()

for (country in countries) {
  prop_cohort_speaks_english_country <- us_df_2020 %>% 
    filter(country_string == country) %>%
    group_by(year_of_immig_groups_string) %>% 
    summarise(
      mean_english_speaker = weighted.mean(english_speaker, perwt, na.rm = TRUE)
    )
  
  # Store the result in the list
  result_list[[country]] <- prop_cohort_speaks_english_country$mean_english_speaker
}

prop_migration_cohort_speaks_english <- data.frame(
  year_of_immigration_groups = prop_cohort_speaks_english_country$year_of_immig_groups_string,
  do.call(cbind, result_list)
) %>%
  mutate(year_of_immigration_groups = factor(year_of_immigration_groups, 
                                             levels = c("Before 1965", "1965 - 1979", "1980 - 1999", "After 1999")))

prop_migration_cohort_speaks_english_long <- prop_migration_cohort_speaks_english %>%
  pivot_longer(cols = -year_of_immigration_groups, 
               names_to = "country", 
               values_to = "proportion")

create_country_plot <- function(country_data) {
  ggplot(country_data, aes(x = proportion, y = year_of_immigration_groups, fill = year_of_immigration_groups)) +
    geom_bar(stat = "identity") +
    geom_text(aes(label = scales::percent(proportion, accuracy = 0.1)), 
              hjust = -0.1, size = 3) +
    scale_x_continuous(labels = scales::percent, limits = c(0, 1)) +
    scale_fill_manual(name = "Cohort",
      values = c("Before 1965" = "lightgreen", 
                 "1965 - 1979" = "red", 
                 "1980 - 1999" = "orange",
                 "After 1999" = "skyblue")) +
    labs(title = str_to_title(unique(country_data$country)),
         x = element_blank(),
         y = element_blank()) +
    theme_cowplot() +
    theme(plot.title = element_text(hjust = 0.5),
          legend.position = "none")  # Remove legend if not needed
}

country_plots <- prop_migration_cohort_speaks_english_long %>%
  split(.$country) %>%
  map(create_country_plot)

combined_plot <- wrap_plots(country_plots, ncol = 2) +
  plot_annotation(
    title = "Proportion of English Speakers by Immigration Cohort and Country of Origin",
    subtitle = "Based on 2020 dataset",
    theme = theme(plot.title = element_text(hjust = 0.5),
                  plot.subtitle = element_text(hjust = 0.5))
  ) &
  theme(legend.position = "bottom")

combined_plot_migration_cohort <- combined_plot + 
  plot_layout(guides = "collect") & 
  theme(legend.position = "bottom")

print(combined_plot_migration_cohort)

ggsave(filename = paste0(out_dir,"migration_cohort_english_speaker_panels.jpg"), plot = combined_plot_migration_cohort, width = 12, height = 9, dpi = 300)
```
# education by cohort migration
```{r}
countries <- c("mexico", "puerto rico", "dominican republic", "cuba")
result_list <- list()

for (country in countries) {
  prop_cohort_education_country <- us_df_2020 %>% 
    filter(country_string == country) %>%
    #filter(!is.na(year_of_immig_groups_string)) %>% 
    group_by(year_of_immig_groups_string) %>% 
    summarise(
      secondary_or_above = weighted.mean(secondary_or_above, perwt, na.rm = TRUE)
    )
  
  # Store the result in the list
  result_list[[country]] <- prop_cohort_education_country$secondary_or_above
}

# Use the year_of_immigration_groups from the last iteration
prop_migration_cohort_education <- data.frame(
  year_of_immigration_groups = prop_cohort_education_country$year_of_immig_groups_string,
  do.call(cbind, result_list)
) %>%
  mutate(year_of_immigration_groups = factor(year_of_immigration_groups, 
                                             levels = c("Before 1965", "1965 - 1979", "1980 - 1999", "After 1999")))

prop_migration_cohort_education_long <- prop_migration_cohort_education %>%
  pivot_longer(cols = -year_of_immigration_groups, 
               names_to = "country", 
               values_to = "proportion")


print(prop_migration_cohort_education_long)

create_country_plot <- function(country_data) {
  ggplot(country_data, aes(x = proportion, y = year_of_immigration_groups, fill = year_of_immigration_groups)) +
    geom_bar(stat = "identity") +
    geom_text(aes(label = scales::percent(proportion, accuracy = 0.1)), 
              hjust = -0.1, size = 3) +
    scale_x_continuous(labels = scales::percent, limits = c(0, 1)) +
    scale_fill_manual(name = "Cohort",
      values = c("Before 1965" = "lightgreen", 
                 "1965 - 1979" = "red", 
                 "1980 - 1999" = "orange",
                 "After 1999" = "skyblue")) +
    labs(title = str_to_title(unique(country_data$country)),
         x = element_blank(),
         y = element_blank()) +
    theme_cowplot() +
    theme(plot.title = element_text(hjust = 0.5),
          legend.position = "none")  # Remove legend if not needed
}

country_plots <- prop_migration_cohort_education_long %>%
  split(.$country) %>%
  map(create_country_plot)

combined_plot <- wrap_plots(country_plots, ncol = 2) +
  plot_annotation(
    title = "Proportion of With At Least Secondary Education by Immigration Cohort and Country of Origin",
    subtitle = "Based on 2020 dataset",
    theme = theme(plot.title = element_text(hjust = 0.5),
                  plot.subtitle = element_text(hjust = 0.5))
  ) &
  theme(legend.position = "bottom")

combined_plot_migration_cohort <- combined_plot + 
  plot_layout(guides = "collect") & 
  theme(legend.position = "bottom")

print(combined_plot_migration_cohort)

ggsave(filename = paste0(out_dir,"migration_cohort_education_panels.jpg"), plot = combined_plot_migration_cohort, width = 12, height = 9, dpi = 300)
```
## are people who migrated at a younger age more likely to get an education at an older age?
```{r}
countries <- c("mexico", "puerto rico", "dominican republic", "cuba")
result_list <- list()

for (country in countries) {
  prop_cohort_education_country <- us_df_2020 %>% 
    filter(country_string == country) %>%
    group_by(age_at_immig_groups_string) %>% 
    summarise(
      secondary_or_above = weighted.mean(secondary_or_above, perwt, na.rm = TRUE)
    )
  
  result_list[[country]] <- prop_cohort_education_country$secondary_or_above
}

prop_age_at_immig_education <- data.frame(
  age_at_immig_groups = prop_cohort_education_country$age_at_immig_groups_string,
  do.call(cbind, result_list)
) %>%
  mutate(age_at_immig_groups = factor(age_at_immig_groups, 
                                      levels = c("Under 15", "15-24", "25-49" ,"50 and above")))

prop_age_at_immig_education_long <- prop_age_at_immig_education %>%
  pivot_longer(cols = -age_at_immig_groups, 
               names_to = "country", 
               values_to = "proportion") %>%
  mutate(country = str_to_title(country))

create_country_plot <- function(country_data) {
  ggplot(country_data, aes(x = proportion, y = age_at_immig_groups, fill = age_at_immig_groups)) +
    geom_bar(stat = "identity") +
    geom_text(aes(label = scales::percent(proportion, accuracy = 0.1)), 
              hjust = -0.1, size = 3) +
    scale_x_continuous(labels = scales::percent, limits = c(0, 1)) +
    scale_fill_manual(name = "Age at Immigration",
      values = c("Under 15" = "#66c2a5", 
                 "15-24" = "#fc8d62",
                 "25-49" = "pink",
                 "50 and above" = "#8da0cb")) +
    labs(title = unique(country_data$country),
         x = "Proportion with At Least Secondary Education",
         y = "Age at Immigration") +
    theme_cowplot() +
    theme(plot.title = element_text(hjust = 0.5),
          legend.position = "none",
          axis.title.x = element_blank())
}

country_plots <- prop_age_at_immig_education_long %>%
  split(.$country) %>%
  map(create_country_plot)

combined_plot <- wrap_plots(country_plots, ncol = 2) +
  plot_annotation(
    title = "Proportion with At Least Secondary Education by Age at Immigration and Country of Origin",
    subtitle = "Based on 2020 dataset",
    theme = theme(plot.title = element_text(hjust = 0.5, size = 14),
                  plot.subtitle = element_text(hjust = 0.5, size = 12))
  ) &
  theme(legend.position = "bottom")

combined_plot_age_at_immig <- combined_plot + 
  plot_layout(guides = "collect") & 
  theme(legend.position = "bottom")

print(combined_plot_age_at_immig)

ggsave(filename = paste0(out_dir,"age_at_immig_education_panels.jpg"), 
       plot = combined_plot_age_at_immig, 
       width = 12, height = 9, dpi = 300)
```
## are people who migrated at a younger age more likely to speak english at an older age?
```{r}
countries <- c("mexico", "puerto rico", "dominican republic", "cuba")
result_list <- list()

for (country in countries) {
  prop_cohort_english_country <- us_df_2020 %>% 
    filter(country_string == country) %>%
    group_by(age_at_immig_groups_string) %>% 
    summarise(
      english_speaker = weighted.mean(english_speaker, perwt, na.rm = TRUE)
    )
  
  result_list[[country]] <- prop_cohort_english_country$english_speaker
}

prop_age_at_immig_english <- data.frame(
  age_at_immig_groups = prop_cohort_english_country$age_at_immig_groups_string,
  do.call(cbind, result_list)
) %>%
  mutate(age_at_immig_groups = factor(age_at_immig_groups, 
                                      levels = c("Under 15", "15-24", "25-49" ,"50 and above")))

prop_age_at_immig_english_long <- prop_age_at_immig_english %>%
  pivot_longer(cols = -age_at_immig_groups, 
               names_to = "country", 
               values_to = "proportion") %>%
  mutate(country = str_to_title(country))

create_country_plot <- function(country_data) {
  ggplot(country_data, aes(x = proportion, y = age_at_immig_groups, fill = age_at_immig_groups)) +
    geom_bar(stat = "identity") +
    geom_text(aes(label = scales::percent(proportion, accuracy = 0.1)), 
              hjust = -0.1, size = 3) +
    scale_x_continuous(labels = scales::percent, limits = c(0, 1)) +
    scale_fill_manual(name = "Age at Immigration",
      values = c("Under 15" = "#66c2a5", 
                 "15-24" = "#fc8d62",
                 "25-49" = "pink",
                 "50 and above" = "#8da0cb")) +
    labs(title = unique(country_data$country),
         x = "Proportion of English Speakers",
         y = "Age at Immigration") +
    theme_cowplot() +
    theme(plot.title = element_text(hjust = 0.5),
          legend.position = "none",
          axis.title.x = element_blank())
}

country_plots <- prop_age_at_immig_english_long %>%
  split(.$country) %>%
  map(create_country_plot)

combined_plot <- wrap_plots(country_plots, ncol = 2) +
  plot_annotation(
    title = "Proportion of English Speakers by Age at Immigration and Country of Origin",
    subtitle = "Based on 2020 dataset",
    theme = theme(plot.title = element_text(hjust = 0.5, size = 14),
                  plot.subtitle = element_text(hjust = 0.5, size = 12))
  ) &
  theme(legend.position = "bottom")

combined_plot_age_at_immig <- combined_plot + 
  plot_layout(guides = "collect") & 
  theme(legend.position = "bottom")

print(combined_plot_age_at_immig)

ggsave(filename = paste0(out_dir,"age_at_immig_english_panels.jpg"), 
       plot = combined_plot_age_at_immig, 
       width = 12, height = 9, dpi = 300)
```
# what proportion of country migrants come from these cohorts?
```{r}
countries <- c("mexico", "puerto rico", "dominican republic", "cuba")
result_list <- list()

for (country in countries) {
  prop_cohort_age_at_immig <- us_df_2020 %>% 
    filter(country_string == country) %>%
    summarise(
      `Under 15` = weighted.mean(age_at_immigration_under15, perwt, na.rm = TRUE),
      `15-24` = weighted.mean(age_at_immigration_15to24, perwt, na.rm = TRUE),
      `25-49` = weighted.mean(age_at_immigration_25to49, perwt, na.rm = TRUE),
      `50 and above` = weighted.mean(age_at_immigration_50plus, perwt, na.rm = TRUE)
    )
  
  result_list[[country]] <- prop_cohort_age_at_immig
}

prop_age_at_immig <- bind_rows(result_list, .id = "country") %>%
  pivot_longer(cols = -country, names_to = "age_at_immig_groups", values_to = "proportion") %>%
  mutate(
    country = str_to_title(country),
    age_at_immig_groups = factor(age_at_immig_groups, levels = c("Under 15", "15-24", "25-49" ,"50 and above"))
  )

create_country_plot <- function(country_data) {
  ggplot(country_data, aes(x = proportion, y = age_at_immig_groups, fill = age_at_immig_groups)) +
    geom_bar(stat = "identity") +
    geom_text(aes(label = scales::percent(proportion, accuracy = 0.1)), 
              hjust = -0.1, size = 3) +
    scale_x_continuous(labels = scales::percent, limits = c(0, 1)) +
    scale_fill_manual(name = "Age at Immigration",
      values = c("Under 15" = "#66c2a5", 
                 "15-24" = "#fc8d62",
                 "25-49" = "pink",
                 "50 and above" = "#8da0cb")) +
    labs(title = unique(country_data$country),
         x = "Proportion of Age at Immigration",
         y = "Age at Immigration") +
    theme_cowplot() +
    theme(plot.title = element_text(hjust = 0.5),
          legend.position = "none",
          axis.title.x = element_blank())
}

country_plots <- prop_age_at_immig %>%
  split(.$country) %>%
  map(create_country_plot)

combined_plot <- wrap_plots(country_plots, ncol = 2) +
  plot_annotation(
    title = "Proportion of Age at Immigration by Country of Origin",
    subtitle = "Based on 2020 dataset",
    theme = theme(plot.title = element_text(hjust = 0.5, size = 14),
                  plot.subtitle = element_text(hjust = 0.5, size = 12))
  ) &
  theme(legend.position = "bottom")

combined_plot_age_at_immig <- combined_plot + 
  plot_layout(guides = "collect") & 
  theme(legend.position = "bottom")

print(combined_plot_age_at_immig)

ggsave(filename = paste0(out_dir,"age_at_immig_prop_panels.jpg"), 
       plot = combined_plot_age_at_immig, 
       width = 12, height = 9, dpi = 300)

```
# percent education within migration cohorts
```{r}
countries <- c("mexico", "puerto rico", "cuba", "dominican republic")
groups <- c("Under 15", "15-24", "25-49", "50 and above")

for (country in countries) {
  result_list <- list()
  
  for (group in groups) {
    prop_cohort_educated <- us_df_2020 %>% 
      filter(country_string == country) %>%
      filter(age_at_immig_groups_string == group) %>%
      summarise(
        less_than_primary = weighted.mean(less_than_primary_completed, perwt, na.rm = TRUE),
        primary = weighted.mean(primary_completed, perwt, na.rm = TRUE),
        secondary = weighted.mean(secondary_completed, perwt, na.rm = TRUE),
        university = weighted.mean(university_completed, perwt, na.rm = TRUE)
      )
    
    result_list[[group]] <- prop_cohort_educated
  }
  
  prop_age_education <- do.call(rbind, result_list) %>%
    rownames_to_column("age_group")
  
  prop_age_education_long <- prop_age_education %>%
    pivot_longer(cols = -age_group, 
                 names_to = "education_level", 
                 values_to = "proportion") %>%
    mutate(education_level = str_to_title(education_level))
  
  ggplot(prop_age_education_long, aes(x = education_level, y = proportion, fill = education_level)) +
    geom_bar(stat = "identity") +
    facet_wrap(~ age_group, scales = "free_y") +
    coord_cartesian(ylim = c(0, 0.75)) +
    theme_minimal() +
    labs(title = paste("Education Levels by Age Group for", str_to_title(country), "Immigrants"),
         x = "Education Level",
         y = "Proportion") +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          legend.position = "none") +
    scale_fill_brewer(palette = "Set2")
  
  ggsave(paste0(plot_dir,"within_migration_cohort_education_levels_", country, ".png"), width = 12, height = 8)
  
}
```

