---
title: "IMR_plot"
author: "Chris Soria"
date: "2025-03-29"
output: html_document
---
#libraries
```{r}
library(cowplot)
library(tidyverse)
library(ggridges)
library(ggpattern)
```
# defining grouped plots
```{r}
create_faceted_age_plot <- function(data, 
                                   age_vars = c("mean_age_at_immigration_15", "mean_age_at_immigration_15to49", "mean_age_at_immigration_50plus"),
                                   var_labels = c("Less than 15", "15 to 49", "50 and above"),
                                   country_var = "country_string",
                                   title = element_blank(),
                                   y_limits = c(0, .7)) {
  
  # Validate inputs
  stopifnot(
    is.data.frame(data),
    all(age_vars %in% names(data)),
    country_var %in% names(data),
    length(var_labels) == length(age_vars)
  )
  
  country_colors <- c(
    "Mexico" = "#E69F00",
    "Puerto Rico" = "#56B4E9",
    "Dominican Republic" = "#009E73",
    "Cuba" = "#CC79A7"
  )
  
  data_long <- data %>%
    pivot_longer(
      cols = all_of(age_vars),
      names_to = "age_group",
      values_to = "mean_age"
    ) %>%
    mutate(age_group = factor(age_group, 
                             levels = age_vars,
                             labels = var_labels))
  
  # Create color mapping
  all_colors <- c()
  for (country in unique(data[[country_var]])) {
    for (j in 1:length(age_vars)) {
      key <- paste0(country, "_", j)
      all_colors[key] <- country_colors[country]
    }
  }
  
  data_long <- data_long %>%
    mutate(fill_group = paste0(.data[[country_var]], "_", as.numeric(age_group)))
  
  data_long[[country_var]] <- factor(
    data_long[[country_var]], 
    levels = names(country_colors)
  )
  
  ggplot(data_long, aes(x = age_group, y = mean_age, fill = fill_group)) +
    geom_bar(stat = "identity", position = "dodge", width = 0.7, color = "black") +
    geom_text(aes(label = round(mean_age, 2)), 
              position = position_dodge(width = 0.7),
              vjust = -0.5,
              size = 4) +
    facet_grid(rows = vars(.data[[country_var]]), switch = "y") +
    scale_fill_manual(values = all_colors) +
    labs(
      title = title,
      x = element_blank(),
      y = element_blank(),
      fill = "Country of Birth"
    ) +
    theme_cowplot() +
    ylim(y_limits) + 
    theme(
      strip.background = element_blank(),
      strip.text.y.left = element_text(size = 12, angle = 0, hjust = 1, face = "bold"),
      axis.text.x = element_text(size = 11, face = "bold"),
      panel.spacing = unit(0.5, "lines"),
      panel.border = element_rect(color = "black", fill = NA, linewidth = 0.5),
      legend.position = "none",
      axis.ticks.y = element_blank(),
      axis.text.y = element_blank(),
      axis.line.y = element_blank(),
      panel.background = element_blank(),
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank()
    )
}


# alternate version of facets where the country is the column
create_faceted_age_plot_alt <- function(data, 
                                   age_vars = c("mean_age_at_immigration_15", "mean_age_at_immigration_15to49", "mean_age_at_immigration_50plus"),
                                   var_labels = c("Less than 15", "15 to 49", "50 and above"),
                                   country_var = "country_string",
                                   title = element_blank(),
                                   y_limits = c(0, .7)) {
  
  # Define specific color mapping for countries
  country_colors <- c(
    "Mexico" = "#E69F00",
    "Puerto Rico" = "#56B4E9",
    "Dominican Republic" = "#009E73",
    "Cuba" = "#CC79A7"
  )
  
  # Reshape data from wide to long format for faceting
  data_long <- data %>%
    pivot_longer(
      cols = all_of(age_vars),
      names_to = "age_group",
      values_to = "mean_age"
    ) %>%
    mutate(age_group = factor(age_group, 
                             levels = age_vars,
                             labels = var_labels))
  
  # Add a combined group for fill mapping
  data_long <- data_long %>%
    mutate(fill_group = paste0(.data[[country_var]], "_", as.numeric(age_group)))
  
  # Create all color combinations for fill_group
  all_colors <- c()
  for (country in unique(data[[country_var]])) {
    for (j in 1:length(age_vars)) {
      key <- paste0(country, "_", j)
      all_colors[key] <- country_colors[country]
    }
  }
  
  # Ensure countries appear in a specific order
  data_long[[country_var]] <- factor(
    data_long[[country_var]], 
    levels = names(country_colors)
  )
  
  ggplot(data_long, aes(x = .data[[country_var]], y = mean_age, fill = fill_group)) +
    geom_bar(stat = "identity", position = "dodge", width = 0.7, color = "black") +
    geom_text(aes(label = round(mean_age, 2)), 
              position = position_dodge(width = 0.7),
              vjust = -0.5,
              size = 4) +
    facet_grid(rows = vars(age_group), switch = "y") +
    scale_fill_manual(values = all_colors) +
    labs(
      title = title,
      x = element_blank(),
      y = element_blank(),
      fill = "Country of Birth"
    ) +
    theme_cowplot() +
    ylim(y_limits) + 
    theme(
      strip.background = element_blank(),
      strip.text.y.left = element_text(size = 12, angle = 0, hjust = 1, face = "bold"),
      axis.text.x = element_text(size = 11, face = "bold", angle = 45, hjust = 1),
      panel.spacing = unit(0.5, "lines"),
      panel.border = element_rect(color = "black", fill = NA, linewidth = 0.5),
      legend.position = "none",
      axis.ticks.y = element_blank(),
      axis.text.y = element_blank(),
      axis.line.y = element_blank(),
      panel.background = element_blank(),
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank()
    )
}

# with four panels
create_combined_faceted_age_plot <- function(data, 
                                           columns = 2,
                                           age_vars = c("mean_age_at_immigration_15", 
                                                       "mean_age_at_immigration_15to49", 
                                                       "mean_age_at_immigration_50plus"),
                                           var_labels = c("Less than 15", "15 to 49", "50 and above"),
                                           country_var = "country_string",
                                           title = element_blank(),
                                           y_limits = c(0, 1)) {
  
  country_colors <- c(
    "Mexico" = "#E69F00",
    "Puerto Rico" = "#56B4E9",
    "Dominican Republic" = "#009E73",
    "Cuba" = "#CC79A7"
  )
  
  data_long <- data %>%
    pivot_longer(
      cols = all_of(age_vars),
      names_to = "age_group",
      values_to = "mean_age"
    ) %>%
    mutate(age_group = factor(age_group, 
                             levels = age_vars,
                             labels = var_labels)) %>%
    mutate(!!sym(country_var) := factor(!!sym(country_var), 
                           levels = c("Mexico", "Puerto Rico", "Dominican Republic", "Cuba"))) %>% 
    arrange(!!sym(country_var))
  
  print(data_long)
  
  countries <- unique(data_long[[country_var]])
  
  plot_list <- lapply(seq_along(countries), function(i) {
    country <- countries[i]
    # Get the specific color for this country
    country_color <- country_colors[country]
    
    # Base plot
    p <- ggplot(data_long[data_long[[country_var]] == country, ], 
           aes(x = age_group, y = mean_age)) +
      geom_bar(stat = "identity", position = "dodge", width = 0.7, 
               color = "black", fill = country_color) +
      geom_text(aes(label = round(mean_age, 2)), 
                position = position_dodge(width = 0.7),
                vjust = -0.5,
                size = 6) +
      labs(title = country,
           x = element_blank(),
           y = element_blank()) +
      theme_cowplot() +
      ylim(y_limits) + 
      theme(
        legend.position = "none",
        axis.text.x = element_text(size = 10, angle = 40, hjust = 1, face = "bold"),
        plot.title = element_text(size = 13, face = "bold", hjust = 0.5),
        panel.grid.major.y = element_line(color = "grey90", size = 0.5)
      )
    
    # Only remove y-axis text for non-first plots when columns=1
    if (columns == 4 && i > 1) {
  p <- p + theme(
    axis.text.y = element_blank(),
    panel.border = element_blank(),
    axis.line.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.line.x = element_line(color = "black")
  )
}
    
    return(p)
  })
  
  combined_plot <- cowplot::plot_grid(plotlist = plot_list, ncol = columns)
  
  # Add the overall title if provided
  if (!identical(title, element_blank())) {
    combined_plot <- cowplot::plot_grid(
      title, combined_plot,
      ncol = 1,
      rel_heights = c(0.1, 1)
    )
  }
  
  return(combined_plot)
}



# simpler age at migration plot function (age groups grouped and country colors)
create_grouped_age_plot <- function(data, 
                                   age_vars = c("mean_age_at_immigration_15", "mean_age_at_immigration_15to49", "mean_age_at_immigration_50plus"),
                                   var_labels = c("Less than 15", "15 to 49", "50 and above"),
                                   country_var = "country_string",
                                   cbPalette, 
                                   title = element_blank(),
                                   y_limits = c(0, .7)) {
  
  # Reshape data from wide to long format
  data_long <- data %>%
    pivot_longer(
      cols = all_of(age_vars),
      names_to = "age_group",
      values_to = "mean_age"
    ) %>%
    mutate(age_group = factor(age_group, 
                             levels = age_vars,
                             labels = var_labels))
  
  # Create the plot with grouped bars
  ggplot(data_long, aes(x = age_group, y = mean_age, fill = .data[[country_var]], group = .data[[country_var]])) +
    geom_bar(stat = "identity", position = position_dodge(width = 0.9), width = 0.8, color = "black") +
    geom_text(aes(label = round(mean_age, 2)), 
              position = position_dodge(width = 0.9),
              vjust = -0.5,
              size = 3.5) +
    scale_fill_manual(values = cbPalette) +
    labs(
      title = title,
      x = element_blank(),
      y = element_blank(),
      fill = "Country of Birth"
    ) +
    theme_cowplot() +
    ylim(y_limits) + 
    theme(
      axis.text.x = element_text(size = 11, face = "bold"),
      panel.border = element_rect(color = "black", fill = NA, linewidth = 0.5),
      legend.position = "bottom",
      legend.title = element_text(face = "bold"),
      panel.background = element_blank(),
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank()
    )
}
```

#defining other plotting functions
```{r}
# for only migrants
create_migrant_plot <- function(data, y_var, 
                                cbPalette, 
                                title = "Proportion of Migrants with at Least a Secondary Degree by Country of Birth",
                                y_limits = c(0,1)) {
  ggplot(data, aes(x = country_string, y = .data[[y_var]], fill = country_string)) +
    geom_bar(stat = "identity", position = "dodge", color = "black", width = 0.7) +
    geom_text(aes(label = round(.data[[y_var]], 2)), # Add labels dynamically
              position = position_dodge(width = 0.9), # Adjust for dodged bars
              vjust = -0.5, # Move text slightly above the bar
              size = 6) + # Adjust text size
    scale_fill_manual(values = cbPalette) +
    labs(
      title = title,
      x = element_blank(),
      y = element_blank(),
      fill = "Country of Origin"
    ) +
    theme_cowplot() +
    ylim(y_limits) + 
    theme(
      axis.text.x = element_text(angle = 40, hjust = 1, size = 14, face = "bold"),
      legend.position = "none",
      panel.grid.major.y = element_line(color = "grey90", size = 0.5)
    )
}


create_migrant_ridges_plot <- function(data, 
                                      value_var,        # Continuous variable (e.g., income, education)
                                      country_var,      # Factor with country names
                                      group_var,        # Migrant vs. non-migrant indicator
                                      cbPalette,        # 2-color palette for groups
                                      title = "Distribution Comparison: Migrants vs Non-Migrants",
                                      y_limits = c(0, 1),
                                      country_order = NULL) {  # Optional country ordering
  
  # Reorder countries if specified
  if(!is.null(country_order)) {
    data <- data %>%
      mutate({{country_var}} := fct_relevel(.data[[country_var]], country_order))
  }

  ggplot(data, aes(
    x = .data[[value_var]], 
    y = .data[[country_var]],
    fill = .data[[group_var]]
  )) +
    geom_density_ridges(
      alpha = 0.6,
      scale = 0.9,
      rel_min_height = 0.01,
      color = "white",
      panel_scaling = FALSE
    ) +
    scale_fill_manual(values = cbPalette) +
    labs(
      title = title,
      x = "Proportion Value",
      y = "Country of Origin",
      fill = "Migration Status"
    ) +
    theme_ridges() +
    coord_cartesian(xlim = y_limits) +
    theme(
      legend.position = "right",
      axis.title.y = element_text(face = "bold", margin = margin(r = 15)),
      axis.text.x = element_text(angle = 40, hjust = 1, size = 13, face = "bold")
    )
}

create_migrant_dotplot <- function(data, 
                                   x_var,          # Continuous variable (e.g., proportion)
                                   group_var,      # Migrant vs. non-migrant column
                                   country_var,    # Country column
                                   cbPalette,      # Palette for countries
                                   title = "Comparison of Migrant and Non-Migrant Groups",
                                   x_limits = c(0, 1),
                                   country_order = NULL) {
  
  # Reorder countries if specified
  if (!is.null(country_order)) {
    data <- data %>%
      mutate({{country_var}} := fct_relevel(.data[[country_var]], country_order))
  }
  
  ggplot(data, aes(
    x = .data[[x_var]], 
    y = 1,  # Dummy y-value for vertical stacking
    shape = .data[[group_var]], 
    color = .data[[country_var]]  # Use country for color
  )) +
    geom_point(size = 5, position = position_dodge(width = 0.5)) +
    # Add connecting red line here
    geom_line(aes(group = .data[[country_var]]), 
              position = position_dodge(width = 0.5), 
              linewidth = 1) +
    facet_grid(rows = vars(.data[[country_var]]), switch = "y") +  # Vertical stacking by country
    scale_shape_manual(values = c(17, 16)) +  # Shapes: Migrant=Circle (16), Non-Migrant=Triangle (17)
    scale_color_manual(values = cbPalette) +  # Consistent colors for countries
    scale_x_continuous(limits = x_limits) +
    labs(
      title = title,
      x = "Proportion Value",
      y = NULL,
      shape = "Migration Status",
      color = "Country of Origin"
    ) +
    theme_cowplot() +
    theme(
      axis.text.y = element_blank(),
      axis.ticks.y = element_blank(),
      panel.grid.major.y = element_blank(),
      strip.background = element_blank(),  # Remove gray background from facet strips
      strip.text.y.left = element_text(size = 12, angle = 0, hjust = 1, face = "bold"), # Make labels larger
      legend.position = "bottom", # Move legend to the bottom
    legend.direction = "horizontal", # Arrange legend items horizontally
    legend.box = "horizontal", # Ensure the box layout is horizontal
    legend.text = element_text(size = 8),# Reduce legend text size
      legend.title = element_text(size = 9),# Reduce legend title size
      panel.spacing = unit(0.5, "lines"),
      panel.border = element_rect(color = "black", fill = NA, linewidth = 0.5)
    )
}
```
# defining parameters
```{r}
countries_of_interest <- c(
  "Puerto Rico",
  "Mexico",
  "Cuba",
  "Dominican Republic"
)

IMR <- read.csv(paste0(data_dir, "UNdata_Export_20250327_063538285.csv")) %>%
  filter(grepl(paste(countries_of_interest, collapse = "|"), Country.or.Area, ignore.case = TRUE)) %>% 
  filter(Country.or.Area != "United States-Mexico-Canada Agreement (USMCA)") %>% 
  filter(Year.s. <2020)

cbPalette <- c("#E69F00", "#56B4E9", "#009E73","#CC79A7")
```
#plot for IMR
```{r}

IMR_plot <- ggplot(IMR, aes(x = Year.s., y = Value, group = Country.or.Area, color = Country.or.Area)) +
  geom_line(size = 1) +  # Adds lines for each country
  labs(
    title = NULL, 
    x = "Year",
    y = "Infant Mortality Rate (per 1,000 births)",
    color = "Country"
  ) +
  theme_cowplot() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = c(0.90, 0.90),
    legend.justification = c("right", "top")
  ) +
  scale_color_manual(values = cbPalette)

ggsave(
    filename = file.path(plot_dir, "IMR_over_time.png"),
    plot = IMR_plot,
    device = "png",
    width = 8,
    height = 6,
    dpi = 300
)
```

#all plots, migrants 2020
```{r}
# with .7 y axis
vars_to_highlight <- c("mean_educ", "mean_lives_alone", "mean_married_cohab", "mean_age_at_immigration_15", "mean_age_at_immigration_25to49")

for (var in vars_to_highlight) {
  current_plot <- create_migrant_plot(
    data = table1,
    y_var = var,
    title = element_blank(),
    cbPalette = cbPalette,
    y_limits = c(0, 0.7) 
  )
  
  ggsave(
    filename = file.path(plot_dir, paste0(var, "_2020_migrants.png")),
    plot = current_plot,
    device = "png",
    width = 9,
    height = 7,
    dpi = 300
  )
}

# with .8 y axis
vars_to_highlight <- c("mean_english_speaker", "mean_is_citizen")

for (var in vars_to_highlight) {
  current_plot <- create_migrant_plot(
    data = table1,
    y_var = var,
    title = element_blank(),
    cbPalette = cbPalette,
    y_limits = c(0, 1) 
  )
  
  ggsave(
    filename = file.path(plot_dir, paste0(var, "_2020_migrants.png")),
    plot = current_plot,
    device = "png",
    width = 12,
    height = 6,
    dpi = 300
  )
}
```
#comparing migrants to residents
```{r}
vars_to_highlight <- c("mean_educ", "mean_lives_alone", "mean_married_cohab")

for (var in vars_to_highlight) {

dot_plot <- create_migrant_dotplot(
  data = combined_table2,
  x_var = var,
  group_var = "group",
  country_var = "country_string",
  cbPalette = cbPalette,
  country_order = c("Mexico", "Puerto Rico", "Dominican Republic", "Cuba"),
  title = element_blank(),
  x_limits = c(0,.75)
)

ggsave(
    filename = file.path(plot_dir, paste0(var, "_2010_migrants_versus_natives.png")),
    plot = dot_plot,
    device = "png",
    width = 12,
    height = 6,
    dpi = 300
  )

}

dot_plot
```
# age at migration plots
```{r}
age_migration_plot <- create_faceted_age_plot(table1,
                                              y_limits = c(0, 1))

age_migration_plot_alt <- create_faceted_age_plot_alt(table1, y_limits = c(0, 1))
create_combined_faceted_age_plot_1 <- create_combined_faceted_age_plot(table1, columns = 2, y_limits = c(0, 1))
create_combined_faceted_age_plot_2 <- create_combined_faceted_age_plot(table1, columns = 4, y_limits = c(0, 1))
age_migration_plot_same_axis <- create_grouped_age_plot(table1, cbPalette = cbPalette, y_limits = c(0, 1))

plot_list <- list(
  v1 = age_migration_plot,
  v2 = age_migration_plot_alt,
  v3 = create_combined_faceted_age_plot_1,
  v4 = create_combined_faceted_age_plot_2,
  v5 = age_migration_plot_same_axis
)

# Loop through the list and save each plot
for (i in names(plot_list)) {
  print(plot_list[[i]])
  ggsave(
    filename = file.path(plot_dir, paste0(i, "_mean_age_migration_2020.png")),
    plot = plot_list[[i]],
    device = "png",
    width = 12,
    height = 6,
    dpi = 300
  )
}
```
# period of immigration plots
```{r}
period_vars <- c("mean_yrimm_before1965", "mean_yrimm_1965_1979", "mean_yrimm_1980_plus")
period_var_labels <- c("Before 1965", "Between 1965 and 1979", "1980 and After")

period_migration_plot <- create_faceted_age_plot(table1,
                                                 age_vars = period_vars,
                                                 var_labels = period_var_labels,
                                              y_limits = c(0, 1))

period_migration_plot_alt <- create_faceted_age_plot_alt(table1, 
                                                         age_vars = period_vars,
                                                         var_labels = period_var_labels,
                                                         y_limits = c(0, 1))

create_combined_faceted_period_plot_1 <- create_combined_faceted_age_plot(table1, 
                                                                          age_vars = period_vars,
                                                                          var_labels = period_var_labels,
                                                                          columns = 2, 
                                                                          y_limits = c(0, 1))

create_combined_faceted_period_plot_2 <- create_combined_faceted_age_plot(table1, 
                                                                          age_vars = period_vars,
                                                                          var_labels = period_var_labels,
                                                                          columns = 4, 
                                                                          y_limits = c(0, 1))

period_migration_plot_same_axis <- create_grouped_age_plot(table1, 
                                                           age_vars = period_vars,
                                                           var_labels = period_var_labels,
                                                           cbPalette = cbPalette, 
                                                           y_limits = c(0, 1))

plot_list <- list(
  v1 = period_migration_plot,
  v2 = period_migration_plot_alt,
  v3 = create_combined_faceted_period_plot_1,
  v4 = create_combined_faceted_period_plot_2,
  v5 = period_migration_plot_same_axis
)

# Loop through the list and save each plot
for (i in names(plot_list)) {
  print(plot_list[[i]])
  ggsave(
    filename = file.path(plot_dir, paste0(i, "_mean_period_migration_2020.png")),
    plot = plot_list[[i]],
    device = "png",
    width = 12,
    height = 6,
    dpi = 300
  )
}
```
# education migrated after 24
```{r}
educ_vars <- c("mean_educ", "mean_educ_24_plus")
educ_var_labels <- c("All Ages", "After Age 24")

educ_24_plus_plot <- create_faceted_age_plot(table1,
                                                 age_vars = educ_vars,
                                                 var_labels = educ_var_labels,
                                              y_limits = c(0, 1))

educ_24_plus_plot_alt <- create_faceted_age_plot_alt(table1, 
                                                         age_vars = educ_vars,
                                                         var_labels = educ_var_labels,
                                                         y_limits = c(0, 1))

create_combined_faceted_educ_24_plus_plot_1 <- create_combined_faceted_age_plot(table1, 
                                                                          age_vars = educ_vars,
                                                         var_labels = educ_var_labels,
                                                                          columns = 2, 
                                                                          y_limits = c(0, 1))

create_combined_faceted_educ_24_plus_plot_2 <- create_combined_faceted_age_plot(table1, 
                                                                          age_vars = educ_vars,
                                                         var_labels = educ_var_labels,
                                                                          columns = 4, 
                                                                          y_limits = c(0, 1))

educ_24_plus_plot_same_axis <- create_grouped_age_plot(table1, 
                                                           age_vars = educ_vars,
                                                         var_labels = educ_var_labels,
                                                           cbPalette = cbPalette, 
                                                           y_limits = c(0, 1))

plot_list <- list(
  v1 = educ_24_plus_plot,
  v2 = educ_24_plus_plot_alt,
  v3 = create_combined_faceted_educ_24_plus_plot_1,
  v4 = create_combined_faceted_educ_24_plus_plot_2,
  v5 = educ_24_plus_plot_same_axis
)

# Loop through the list and save each plot
for (i in names(plot_list)) {
  print(plot_list[[i]])
  ggsave(
    filename = file.path(plot_dir, paste0(i, "_educ_24_plus_2020.png")),
    plot = plot_list[[i]],
    device = "png",
    width = 12,
    height = 6,
    dpi = 300
  )
}
```
#education, international
```{r}
educ_vars <- c("mean_educ", "mean_educ_international")
educ_var_labels <- c("US Migrants", "Native Country Residents")

educ_international_plot <- create_faceted_age_plot(table2,
                                                 age_vars = educ_vars,
                                                 var_labels = educ_var_labels,
                                              y_limits = c(0, 1))

educ_international_plot_alt <- create_faceted_age_plot_alt(table2, 
                                                         age_vars = educ_vars,
                                                         var_labels = educ_var_labels,
                                                         y_limits = c(0, 1))

create_combined_faceted_educ_international_plot_1 <- create_combined_faceted_age_plot(table2, 
                                                                          age_vars = educ_vars,
                                                         var_labels = educ_var_labels,
                                                                          columns = 2, 
                                                                          y_limits = c(0, 1))

create_combined_faceted_educ_international_plot_2 <- create_combined_faceted_age_plot(table2, 
                                                                          age_vars = educ_vars,
                                                         var_labels = educ_var_labels,
                                                                          columns = 4, 
                                                                          y_limits = c(0, 1))

educ_international_plot_same_axis <- create_grouped_age_plot(table2, 
                                                           age_vars = educ_vars,
                                                         var_labels = educ_var_labels,
                                                           cbPalette = cbPalette, 
                                                           y_limits = c(0, 1))

plot_list <- list(
  v1 = educ_international_plot,
  v2 = educ_international_plot_alt,
  v3 = create_combined_faceted_educ_international_plot_1,
  v4 = create_combined_faceted_educ_international_plot_2,
  v5 = educ_international_plot_same_axis
)

# Loop through the list and save each plot
for (i in names(plot_list)) {
  print(plot_list[[i]])
  ggsave(
    filename = file.path(plot_dir, paste0(i, "_educ_international_comp_2010.png")),
    plot = plot_list[[i]],
    device = "png",
    width = 12,
    height = 6,
    dpi = 300
  )
}
```
#lives_alone, international
```{r}
lives_alone_vars <- c("mean_lives_alone", "mean_lives_alone_international")
lives_alone_var_labels <- c("US Migrants", "Native Country Residents")

lives_alone_international_plot <- create_faceted_age_plot(table2,
                                                 age_vars = lives_alone_vars,
                                                 var_labels = lives_alone_var_labels,
                                              y_limits = c(0, .5))

lives_alone_international_plot_alt <- create_faceted_age_plot_alt(table2, 
                                                         age_vars = lives_alone_vars,
                                                         var_labels = lives_alone_var_labels,
                                                         y_limits = c(0, .5))

create_combined_faceted_lives_alone_international_plot_1 <- create_combined_faceted_age_plot(table2, 
                                                                          age_vars = lives_alone_vars,
                                                         var_labels = lives_alone_var_labels,
                                                                          columns = 2, 
                                                                          y_limits = c(0, .5))

create_combined_faceted_lives_alone_international_plot_2 <- create_combined_faceted_age_plot(table2, 
                                                                          age_vars = lives_alone_vars,
                                                         var_labels = lives_alone_var_labels,
                                                                          columns = 4, 
                                                                          y_limits = c(0, .5))

lives_alone_international_plot_same_axis <- create_grouped_age_plot(table2, 
                                                           age_vars = lives_alone_vars,
                                                         var_labels = lives_alone_var_labels,
                                                           cbPalette = cbPalette, 
                                                           y_limits = c(0, .5))

plot_list <- list(
  v1 = lives_alone_international_plot,
  v2 = lives_alone_international_plot_alt,
  v3 = create_combined_faceted_lives_alone_international_plot_1,
  v4 = create_combined_faceted_lives_alone_international_plot_2,
  v5 = lives_alone_international_plot_same_axis
)

# Loop through the list and save each plot
for (i in names(plot_list)) {
  print(plot_list[[i]])
  ggsave(
    filename = file.path(plot_dir, paste0(i, "_lives_alone_international_comp_2010.png")),
    plot = plot_list[[i]],
    device = "png",
    width = 12,
    height = 6,
    dpi = 300
  )
}
```
#marr_cohab, international
```{r}
married_cohab_vars <- c("mean_married_cohab", "mean_married_cohab_international")
married_cohab_var_labels <- c("US Migrants", "Native Country Residents")

married_cohab_international_plot <- create_faceted_age_plot(table2,
                                                 age_vars = married_cohab_vars,
                                                 var_labels = married_cohab_var_labels,
                                              y_limits = c(0, 1))

married_cohab_international_plot_alt <- create_faceted_age_plot_alt(table2, 
                                                         age_vars = married_cohab_vars,
                                                         var_labels = married_cohab_var_labels,
                                                         y_limits = c(0, 1))

create_combined_faceted_married_cohab_international_plot_1 <- create_combined_faceted_age_plot(table2, 
                                                                          age_vars = married_cohab_vars,
                                                         var_labels = married_cohab_var_labels,
                                                                          columns = 2, 
                                                                          y_limits = c(0, 1))

create_combined_faceted_married_cohab_international_plot_2 <- create_combined_faceted_age_plot(table2, 
                                                                          age_vars = married_cohab_vars,
                                                         var_labels = married_cohab_var_labels,
                                                                          columns = 4, 
                                                                          y_limits = c(0, 1))

married_cohab_international_plot_same_axis <- create_grouped_age_plot(table2, 
                                                           age_vars = married_cohab_vars,
                                                         var_labels = married_cohab_var_labels,
                                                           cbPalette = cbPalette, 
                                                           y_limits = c(0, 1))

plot_list <- list(
  v1 = married_cohab_international_plot,
  v2 = married_cohab_international_plot_alt,
  v3 = create_combined_faceted_married_cohab_international_plot_1,
  v4 = create_combined_faceted_married_cohab_international_plot_2,
  v5 = married_cohab_international_plot_same_axis
)

# Loop through the list and save each plot
for (i in names(plot_list)) {
  print(plot_list[[i]])
  ggsave(
    filename = file.path(plot_dir, paste0(i, "_married_cohab_international_comp_2010.png")),
    plot = plot_list[[i]],
    device = "png",
    width = 12,
    height = 6,
    dpi = 300
  )
}
```
#education, migrants, 2010/2020
```{r}
educ_vars <- c("mean_educ", "mean_educ_2020")
educ_var_labels <- c("2010", "2020")

educ_2010_2020_plot <- create_faceted_age_plot(table2,
                                                 age_vars = educ_vars,
                                                 var_labels = educ_var_labels,
                                              y_limits = c(0, 1))

educ_2010_2020_plot_alt <- create_faceted_age_plot_alt(table2, 
                                                         age_vars = educ_vars,
                                                 var_labels = educ_var_labels,
                                                         y_limits = c(0, 1))

create_combined_faceted_educ_2010_2020_plot_1 <- create_combined_faceted_age_plot(table2, 
                                                                          age_vars = educ_vars,
                                                 var_labels = educ_var_labels,
                                                                          columns = 2, 
                                                                          y_limits = c(0, 1))

create_combined_faceted_educ_2010_2020_plot_2 <- create_combined_faceted_age_plot(table2, 
                                                                          age_vars = educ_vars,
                                                 var_labels = educ_var_labels,
                                                                          columns = 4, 
                                                                          y_limits = c(0, 1))

educ_2010_2020_plot_same_axis <- create_grouped_age_plot(table2, 
                                                           age_vars = educ_vars,
                                                 var_labels = educ_var_labels,
                                                           cbPalette = cbPalette, 
                                                           y_limits = c(0, 1))

plot_list <- list(
  v1 = educ_2010_2020_plot,
  v2 = educ_2010_2020_plot_alt,
  v3 = create_combined_faceted_educ_2010_2020_plot_1,
  v4 = create_combined_faceted_educ_2010_2020_plot_2,
  v5 = educ_2010_2020_plot_same_axis
)

# Loop through the list and save each plot
for (i in names(plot_list)) {
  print(plot_list[[i]])
  ggsave(
    filename = file.path(plot_dir, paste0(i, "_educ_comp_2010_2020.png")),
    plot = plot_list[[i]],
    device = "png",
    width = 12,
    height = 6,
    dpi = 300
  )
}
```

# source of migrants
```{r}
background_table <- background_table %>% 
  filter(Country %in% c("Mexico", "Puerto Rico", "Cuba", "Dominican Republic"))

percent_migration <- ggplot(background_table, aes(x = reorder(Country, -`Percent of Migrants`), y = `Percent of Migrants`, fill = Country)) +
  geom_bar(stat = "identity") +
  labs(
    title = "Percent of Older Hispanic Migrants (60+) to the US by Country of Origin",
    x = element_blank(),
    y = element_blank()
  ) +
  theme_cowplot() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "none"
  ) +
  scale_x_discrete(labels = function(x) ifelse(seq_along(x) <= 4, x, ""))

ggsave(
    filename = file.path(plot_dir, "percent_migration_2020.png"),
    plot = percent_migration,
    device = "png",
    width = 12,
    height = 6,
    dpi = 300
  )
```
# stress data (from Cervantez et al.)
```{r}
stress_data <- data.frame(
  Group = c("Mexican", "Puerto Rican", "Cuban", "Central American", 
            "South American", "Dominican"),
  Parental = c(17.78, 14.25, 15.65, 14.66, 15.62, 14.83),
  Parental_SD = c(8.60, 3.93, 7.61, 5.28, 5.78, 4.25),
  Marital = c(15.83, 14.28, 12.95, 13.54, 15.13, 13.22),
  Marital_SD = c(7.98, 5.66, 4.49, 6.26, 7.43, 4.53),
  Occupational = c(18.11, 16.14, 16.94, 16.46, 19.44, 17.22),
  Occupational_SD = c(8.45, 6.19, 8.32, 6.67, 9.58, 6.82),
  Discrimination = c(12.87, 12.37, 11.34, 12.35, 12.51, 12.11),
  Discrimination_SD = c(5.24, 4.46, 3.94, 4.85, 5.78, 3.93),
  Immigration = c(13.41, 10.38, 9.75, 10.68, 11.43, 9.98),
  Immigration_SD = c(7.43, 3.60, 2.23, 4.20, 5.62, 2.55),
  Marital_Acculturation = c(11.09, 10.06, 10.10, 9.45, 10.57, 9.78),
  Marital_Acculturation_SD = c(4.68, 2.92, 2.98, 1.56, 3.13, 2.20),
  Health = c(13.23, 9.35, 11.14, 10.18, 13.05, 9.87),
  Health_SD = c(7.08, 2.25, 5.42, 4.35, 7.79, 4.69),
  Language = c(9.97, 6.95, 8.89, 8.99, 8.82, 11.42),
  Language_SD = c(5.21, 2.16, 4.93, 4.88, 4.43, 5.55),
  Premigration = c(9.80, 6.92, 8.44, 8.28, 8.12, 9.17),
  Premigration_SD = c(5.92, 1.89, 4.16, 4.27, 4.04, 4.30),
  Family = c(7.93, 6.50, 6.78, 7.03, 7.74, 6.72),
  Family_SD = c(3.95, 2.50, 3.57, 2.88, 4.09, 2.37),
  Total = c(115.17, 89.29, 97.95, 100.28, 111.94, 99.10),
  Total_SD = c(44.15, 15.29, 32.28, 25.24, 39.93, 18.73)
)

ggplot(stress_data, aes(x = Group, y = Health)) +
  geom_point(color = "black", size = 3) +
  #geom_errorbar(aes(ymin = Health - Health_SD, ymax = Health + Health_SD), 
                #width = 0.2, color = "black") +
  theme_light() +
  labs(title = "Self-Rated Health Stress by Hispanic/Latino Group",
       x = "Country of Birth",
       y = "Health Stress") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 12))
```

