---
title: "US International SES - Data Preprocessing"
author: "Chris Soria"
date: "2025-01-08"
output: html_document
---

# Overview

This script processes US Census/ACS data and international census data for comparative analysis of older adult (60+) Hispanic immigrants.

**Outputs:**
- `data/US_2010_v100.rds` (ACS 2010)
- `data/US_2020_v100.rds` (ACS 2020)
- `data/US_2010_v100_census.rds` (Census 2010 - International IPUMS format)
- `data/CuDrMePrUs_10_12.rds` (Combined international)

# Setup

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(haven)
library(labelled)

# Set working directory
target_dir <- normalizePath("/Users/chrissoria/Documents/Research/us_international_ses")
current_dir <- normalizePath(getwd())

if (grepl("us_international_ses", current_dir)) {
  data_dir <- file.path("/Users/chrissoria/Documents/Research/us_international_ses/data/")
  weights_dir <- file.path("/Users/chrissoria/Documents/Research/us_international_ses/data/weights/")
} else {
  data_dir <- file.path("C:/Users/Ty/Desktop/ses_international/data/")
  weights_dir <- file.path("C:/Users/Ty/Desktop/ses_international/data/weights/")
}

# Create weights directory if it doesn't exist
dir.create(weights_dir, showWarnings = FALSE, recursive = TRUE)
```

# Helper Functions

```{r helper_functions}
# Function: Create age group variables
create_age_groups <- function(df) {
  df %>%
    mutate(
      age_groups = case_when(
        age < 70 ~ 1,
        age >= 70 & age < 80 ~ 2,
        age >= 80 & age < 90 ~ 3,
        age >= 90 ~ 4
      ),
      age_groups = factor(age_groups, levels = 1:4,
                          labels = c("60-69", "70-79", "80-89", "90+")),
      age_60to69 = as.integer(age >= 60 & age < 70),
      age_70to79 = as.integer(age >= 70 & age < 80),
      age_80to89 = as.integer(age >= 80 & age < 90),
      age_90plus = as.integer(age >= 90)
    )
}

# Function: Create year of immigration variables
# yrimm_var should be the name of the year of immigration column
create_yrimm_vars <- function(df, yrimm_var) {
  yrimm_sym <- sym(yrimm_var)

  df %>%
    mutate(
      year_of_immigration_groups = case_when(
        !!yrimm_sym < 1965 & !!yrimm_sym != 0 ~ 1,
        !!yrimm_sym >= 1965 & !!yrimm_sym < 1981 & !!yrimm_sym != 0 ~ 2,
        !!yrimm_sym >= 1980 & !!yrimm_sym < 2000 & !!yrimm_sym != 0 ~ 3,
        !!yrimm_sym >= 2000 & !!yrimm_sym != 0 ~ 4
      ),
      year_of_immigration_groups = factor(year_of_immigration_groups, levels = 1:4,
                                          labels = c("Before 1965", "1965 - 1979",
                                                    "1980 - 1999", "After 1999")),
      year_of_immig_groups_string = as.character(year_of_immigration_groups),
      yrimm_before1965 = as.integer(!!yrimm_sym < 1965 & !!yrimm_sym != 0),
      yrimm_1965to1980 = as.integer(!!yrimm_sym >= 1965 & !!yrimm_sym < 1981 & !!yrimm_sym != 0),
      yrimm_1980to1999 = as.integer(!!yrimm_sym >= 1980 & !!yrimm_sym < 2000 & !!yrimm_sym != 0),
      yrimm_2000plus = as.integer(!!yrimm_sym >= 2000 & !!yrimm_sym != 0)
    )
}

# Function: Create age at immigration variables
create_age_at_immig_vars <- function(df) {
  df %>%
    mutate(
      age_at_immigration_groups = case_when(
        age_at_immigration < 15 ~ 1,
        age_at_immigration >= 15 & age_at_immigration < 25 ~ 2,
        age_at_immigration >= 25 & age_at_immigration < 50 ~ 3,
        age_at_immigration >= 50 ~ 4
      ),
      age_at_immigration_groups = factor(age_at_immigration_groups, levels = 1:4,
                                         labels = c("Under 15", "15-24", "25-49", "50 and above")),
      age_at_immig_groups_string = as.character(age_at_immigration_groups),
      age_at_immigration_under15 = as.integer(age_at_immigration < 15),
      age_at_immigration_15to24 = as.integer(age_at_immigration >= 15 & age_at_immigration < 25),
      age_at_immigration_25to49 = as.integer(age_at_immigration >= 25 & age_at_immigration < 50),
      age_at_immigration_50plus = as.integer(age_at_immigration >= 50),
      age_at_immigration_15to49 = as.integer(age_at_immigration >= 15 & age_at_immigration < 50)
    )
}

# Function: Create race/nativity category (12-category classification)
create_race_nativity_category <- function(df) {
  df %>%
    mutate(
      race_native_category = case_when(
        native_foreign_race == "foreign-born black hispanic" ~ 1,
        native_foreign_race == "foreign-born white hispanic" ~ 2,
        native_foreign_race == "foreign-born other hispanic" ~ 3,
        native_foreign_race == "foreign-born black not hispanic" ~ 4,
        native_foreign_race == "foreign-born white not hispanic" ~ 5,
        native_foreign_race == "foreign-born other not hispanic" ~ 6,
        native_foreign_race == "native-born black hispanic" ~ 7,
        native_foreign_race == "native-born white hispanic" ~ 8,
        native_foreign_race == "native-born other hispanic" ~ 9,
        native_foreign_race == "native-born black not hispanic" ~ 10,
        native_foreign_race == "native-born white not hispanic" ~ 11,
        native_foreign_race == "native-born other not hispanic" ~ 12
      ),
      race_native_category = factor(race_native_category, levels = 1:12,
                                    labels = c("Foreign-Born Black Hispanic",
                                              "Foreign-Born White Hispanic",
                                              "Foreign-Born Other Hispanic",
                                              "Foreign-Born Black Non-Hispanic",
                                              "Foreign-Born White Non-Hispanic",
                                              "Foreign-Born Other Non-Hispanic",
                                              "Native-Born Black Hispanic",
                                              "Native-Born White Hispanic",
                                              "Native-Born Other Hispanic",
                                              "Native-Born Black Non-Hispanic",
                                              "Native-Born White Non-Hispanic",
                                              "Native-Born Other Non-Hispanic"))
    )
}

# Function: Create education variables (for ACS detailed codes)
create_education_vars_acs <- function(df) {
  df %>%
    mutate(
      edattain_string = tolower(as.character(as_factor(educd))),
      less_than_primary_completed = as.integer(edattain_string %in% c(
        "grade 1", "grade 2", "grade 3", "grade 4", "grade 5",
        "kindergarten", "no schooling completed", "nursery school, preschool"
      )),
      primary_completed = as.integer(edattain_string %in% c(
        "grade 6", "grade 7", "grade 8", "grade 9", "grade 10", "grade 11"
      )),
      secondary_completed = as.integer(edattain_string %in% c(
        "regular high school diploma", "ged or alternative credential",
        "1 or more years of college credit, no degree",
        "some college, but less than 1 year", "12th grade, no diploma",
        "associate's degree, type not specified"
      )),
      university_completed = as.integer(edattain_string %in% c(
        "bachelor's degree", "master's degree", "doctoral degree",
        "professional degree beyond a bachelor's degree"
      )),
      primary_or_above = as.integer(primary_completed == 1 | secondary_completed == 1 | university_completed == 1),
      secondary_or_above = as.integer(secondary_completed == 1 | university_completed == 1)
    )
}

# Function: Create education variables (for international census)
create_education_vars_intl <- function(df) {
  df %>%
    mutate(
      edattain_string = as.character(as_factor(edattain)),
      less_than_primary_completed = as.integer(edattain_string == "less than primary completed"),
      primary_completed = as.integer(edattain_string == "primary completed"),
      secondary_completed = as.integer(edattain_string == "secondary completed"),
      university_completed = as.integer(edattain_string == "university completed"),
      primary_or_above = as.integer(primary_completed == 1 | secondary_completed == 1 | university_completed == 1),
      secondary_or_above = as.integer(secondary_completed == 1 | university_completed == 1)
    )
}

# Function: Create household variables
create_household_vars <- function(df) {
  df %>%
    mutate(
      lives_alone = as.integer(famsize == 1 & !is.na(famsize)),
      child_in_household = as.integer(nchild > 0 & !is.na(nchild))
    )
}

# Function: Create sex dummy variables
create_sex_vars <- function(df) {
  df %>%
    mutate(
      male = as.integer(sex == 1),
      female = as.integer(sex == 2)
    )
}

# US states list for nativity detection
us_states <- c("alabama", "alaska", "arizona", "arkansas", "california", "colorado",
               "connecticut", "delaware", "florida", "georgia", "hawaii", "idaho",
               "illinois", "indiana", "iowa", "kansas", "kentucky", "louisiana",
               "maine", "maryland", "massachusetts", "michigan", "minnesota",
               "mississippi", "missouri", "montana", "nebraska", "nevada",
               "new hampshire", "new jersey", "new mexico", "new york",
               "north carolina", "north dakota", "ohio", "oklahoma", "oregon",
               "pennsylvania", "rhode island", "south carolina", "south dakota",
               "tennessee", "texas", "utah", "vermont", "virginia", "washington",
               "west virginia", "wisconsin", "wyoming", "district of columbia")
```

# Part 1: Process US ACS Data (2010 and 2020)

```{r process_acs, message=FALSE, warning=FALSE}
process_acs_data <- function(data_file, sample_code, year_val) {

  df <- read_dta(paste0(data_dir, data_file)) %>%
    filter(sample == sample_code) %>%
    filter(age > 59)

  # Hispanic Category Classification
  df <- df %>%
    mutate(
      HispanicCategory = case_when(
        bpl == 210 & hispan != 0 ~ 3,  # Central American
        (bpl < 400 | bpl > 599) & hispan != 0 & bpl != 210 ~ 2,  # Other Latino
        (bpl >= 400 & bpl <= 599) & hispan != 0 ~ 1,  # Hispanic Non-Latino
        TRUE ~ 0  # Not Latino
      ),
      HispanicCategory = factor(HispanicCategory, levels = 0:3,
                                labels = c("Not Latino", "Hispanic Non-Latino",
                                          "Other Latino", "Central-American Latino")),
      HispanicCategoryString = as.character(HispanicCategory)
    )

  # Create derived variables
  df <- df %>%
    create_household_vars() %>%
    create_age_groups() %>%
    create_yrimm_vars("yrimmig")

  # Nativity: Identify US-born via state of birth
  df <- df %>%
    mutate(
      bpl_string = tolower(as.character(as_factor(bpl))),
      contains_us_state = as.integer(sapply(bpl_string, function(x) {
        any(sapply(us_states, function(state) grepl(state, x, fixed = TRUE)))
      })),
      bplcountry = as.character(as_factor(bpld)),
      bplcountry = if_else(contains_us_state == 1, "united states", tolower(bplcountry)),
      nativity_string = if_else(contains_us_state == 1, "native-born", "foreign-born")
    )

  # Race string processing
  df <- df %>%
    mutate(
      race_string = tolower(as.character(as_factor(race))),
      race_string = case_when(
        race_string == "black/african american" ~ "black",
        race_string %in% c("white", "black") ~ race_string,
        TRUE ~ "other"
      ),
      hispan_string = tolower(as.character(as_factor(hispan))),
      hispan_string = if_else(hispan_string != "not hispanic", "hispanic", hispan_string),
      # Fix: Dominicans often misclassified
      hispan_string = if_else(bpld == 26010, "hispanic", hispan_string),
      race_string = paste(race_string, hispan_string),
      native_foreign_race = paste(nativity_string, race_string)
    )

  df <- df %>% create_race_nativity_category()

  # Marital status
  df <- df %>%
    mutate(
      married_cohab = as.integer(marst %in% c(1, 2) & !is.na(marst)),
      married_not_cohab = as.integer(marst == 2 & !is.na(marst))
    )

  # Immigration timing
  df <- df %>%
    mutate(
      years_in_us = yrsusa1,
      age_at_immigration = age - years_in_us,
      age_at_immigration = if_else(nativity_string == "native-born", NA_real_, age_at_immigration)
    ) %>%
    create_age_at_immig_vars()

  # Citizenship
  df <- df %>%
    mutate(
      is_naturalized_citizen = as.integer(citizen == 2),
      is_born_citizen = as.integer(citizen == 0),
      is_citizen = as.integer(citizen != 3)
    )

  df <- df %>% create_sex_vars()

  # English proficiency
  df <- df %>%
    mutate(
      speakeng_string = tolower(as.character(as_factor(speakeng))),
      english_speaker = as.integer(speakeng_string != "does not speak english"),
      english_speaker_not_well = as.integer(speakeng_string == "yes, but not well"),
      english_speaker_well = as.integer(english_speaker == 1 & english_speaker_not_well == 0)
    )

  df <- df %>% create_education_vars_acs()

  # Country-year identifiers
  df <- df %>%
    mutate(
      year_str = as.character(year),
      country_string = tolower(bplcountry),
      country_year = paste0(country_string, "_", year_str),
      hispanic_migrant_status = paste(nativity_string, hispan_string),
      hispanic_category_migrant_status = paste(nativity_string, HispanicCategoryString),
      hispanic_category_migrant_status = if_else(
        hispanic_category_migrant_status == "foreign-born Hispanic Non-Latino",
        "foreign-born Not Latino",
        hispanic_category_migrant_status
      ),
      hispanic_migrant_status_race = paste(nativity_string, race_string)
    )

  # Add cognitive/independence difficulty variables if they exist
  if ("diffrem" %in% names(df)) {
    df <- df %>%
      mutate(
        cognitive_difficulty = as.integer(diffrem == 2),
        independent_living_difficulty = as.integer(diffcare == 2)
      )
  }

  df
}

# Process 2020 ACS (both use usa_00007.dta, filtered by sample code)
us_df_2020 <- process_acs_data("usa_00007.dta", 202003, 2020)
saveRDS(us_df_2020, paste0(data_dir, "US_2020_v100.rds"))
write_dta(us_df_2020, paste0(data_dir, "US_2020_v100.dta"))
cat("Saved US 2020 ACS data:", nrow(us_df_2020), "observations\n")

# Process 2010 ACS
us_df_2010 <- process_acs_data("usa_00007.dta", 201003, 2010)
saveRDS(us_df_2010, paste0(data_dir, "US_2010_v100.rds"))
write_dta(us_df_2010, paste0(data_dir, "US_2010_v100.dta"))
cat("Saved US 2010 ACS data:", nrow(us_df_2010), "observations\n")

rm(us_df_2010, us_df_2020)
gc()
```

# Part 2: Process US Census 2010 (International IPUMS format)

```{r process_census_2010, message=FALSE, warning=FALSE}
us_census_2010 <- read_dta(paste0(data_dir, "ipumsi_00002_US_2010.dta")) %>%
  filter(age > 59)

# Hispanic Category (using international bplcountry codes)
# Central America: 22010-22080 (Belize, Costa Rica, El Salvador, Guatemala, Honduras, Nicaragua, Panama)
us_census_2010 <- us_census_2010 %>%
  mutate(
    HispanicCategory = case_when(
      (bplcountry >= 22010 & bplcountry <= 22080) & hispan != 0 ~ 3,
      hispan != 0 & !((bplcountry >= 22010 & bplcountry <= 22080)) ~ 2,
      TRUE ~ 0
    ),
    HispanicCategory = if_else(HispanicCategory == 0 & hispan == 0, 0, HispanicCategory),
    HispanicCategory = factor(HispanicCategory, levels = 0:3,
                              labels = c("Not Latino", "Hispanic Non-Latino",
                                        "Other Latino", "Central-American Latino")),
    HispanicCategoryString = as.character(HispanicCategory)
  )

us_census_2010 <- us_census_2010 %>%
  create_household_vars()

# Latin American identifier
us_census_2010 <- us_census_2010 %>%
  mutate(
    latin_american = as.integer(bpl > 21019 & bpl < 24000),
    latin_american = if_else(hispan > 0 & bpl == 24040, 1L, latin_american),
    hispan = if_else(latin_american == 0, 0, hispan)
  )

# Calculate age standardization weights (by sex)
us_census_2010 <- us_census_2010 %>%
  mutate(
    sex_string = tolower(as.character(as_factor(sex))),
    age2_string = as.character(as_factor(age2))
  )

# Calculate US age weights by sex
us_age_weights <- us_census_2010 %>%
  group_by(sex_string, age2, age2_string) %>%
  summarise(perwt_sum = sum(perwt, na.rm = TRUE), .groups = "drop") %>%
  group_by(sex_string) %>%
  mutate(
    total_perwt = sum(perwt_sum),
    age_weight = perwt_sum / total_perwt,
    match_var_us = paste0(sex_string, age2_string)
  ) %>%
  ungroup() %>%
  select(sex_string, age2, age2_string, age_weight, match_var_us)

# Save weights
for (s in c("female", "male")) {
  us_age_weights %>%
    filter(sex_string == s) %>%
    rename(!!paste0("age_weight_", s) := age_weight) %>%
    write_csv(paste0(weights_dir, "us_age_weights_", s, ".csv"))
}

# Continue with variable creation
us_census_2010 <- us_census_2010 %>%
  mutate(country_string = as.character(as_factor(bplcountry))) %>%
  create_age_groups() %>%
  create_yrimm_vars("yrimm")

# Nativity and race
us_census_2010 <- us_census_2010 %>%
  mutate(
    nativity_string = as.character(as_factor(nativity)),
    race_string = as.character(as_factor(race)),
    hispan_string = as.character(as_factor(hispan)),
    hispan_string = if_else(hispan_string != "not hispanic", "hispanic", hispan_string),
    hispan_string = if_else(bpl == 26010, "hispanic", hispan_string),  # Fix Dominican
    race_string = if_else(!race_string %in% c("white", "black"), "other", race_string),
    race_string = paste(race_string, hispan_string),
    native_foreign_race = paste(nativity_string, race_string)
  ) %>%
  create_race_nativity_category()

# Marital status (international coding: 2 = married/cohabiting)
us_census_2010 <- us_census_2010 %>%
  mutate(married_cohab = as.integer(marst == 2 & !is.na(marst)))

# Immigration timing
us_census_2010 <- us_census_2010 %>%
  mutate(
    years_in_us = 2010 - yrimm,
    years_in_us = if_else(years_in_us == 2010, NA_real_, years_in_us),
    age_at_immigration = age - years_in_us
  ) %>%
  create_age_at_immig_vars()

# Citizenship (international coding: 3 = naturalized, 2 = birth citizen)
us_census_2010 <- us_census_2010 %>%
  mutate(
    is_naturalized_citizen = as.integer(citizen == 3),
    is_birth_citizen = as.integer(citizen == 2)
  ) %>%
  create_sex_vars()

# English proficiency
us_census_2010 <- us_census_2010 %>%
  mutate(english_speaker = as.integer(speakeng == 1))

# Education with years of schooling correction
us_census_2010 <- us_census_2010 %>%
  mutate(edattain_string = as.character(as_factor(edattain)))

# Correct education based on years of schooling
us_census_2010 <- us_census_2010 %>%
  mutate(
    edattain = case_when(
      yrschool < 12 & edattain_string == "secondary completed" ~ 2,
      yrschool > 12 & edattain_string == "primary completed" ~ 3,
      yrschool > 15 & edattain_string == "secondary completed" ~ 4,
      TRUE ~ edattain
    )
  ) %>%
  select(-edattain_string) %>%
  create_education_vars_intl()

# Country-year identifiers
us_census_2010 <- us_census_2010 %>%
  mutate(
    year_str = as.character(year),
    country_year = paste0(country_string, "_", year_str),
    hispanic_migrant_status = paste(nativity_string, hispan_string),
    hispanic_category_migrant_status = paste(nativity_string, HispanicCategoryString),
    hispanic_category_migrant_status = if_else(
      hispanic_category_migrant_status == "foreign-born Hispanic Non-Latino",
      "foreign-born Not Latino",
      hispanic_category_migrant_status
    ),
    hispanic_migrant_status_race = paste(nativity_string, race_string)
  )

saveRDS(us_census_2010, paste0(data_dir, "US_2010_v100_census.rds"))
write_dta(us_census_2010, paste0(data_dir, "US_2010_v100_census.dta"))
cat("Saved US Census 2010 data:", nrow(us_census_2010), "observations\n")

rm(us_census_2010)
gc()
```

# Part 3: Process International Census Data with Age Standardization

```{r process_international, message=FALSE, warning=FALSE}
# Load international data
intl_df <- read_dta(paste0(data_dir, "ses_v2.dta"))

# Append US 2010 census data
us_census_2010 <- read_dta(paste0(data_dir, "ipumsi_00002_US_2010.dta"))
intl_df <- bind_rows(intl_df, us_census_2010)
rm(us_census_2010)

intl_df <- intl_df %>%
  mutate(country_string = as.character(as_factor(country))) %>%
  filter(age > 59)

intl_df <- intl_df %>%
  create_household_vars() %>%
  create_age_groups() %>%
  create_sex_vars()

# Education
intl_df <- intl_df %>%
  mutate(
    edattain_string = as.character(as_factor(edattain)),
    edattaind_string = as.character(as_factor(edattaind)),
    less_than_primary_completed = as.integer(edattain_string == "less than primary completed"),
    primary_completed = as.integer(edattain_string == "primary completed"),
    secondary_completed = as.integer(edattain_string == "secondary completed"),
    university_completed = as.integer(edattain_string == "university completed"),
    education_unknown = as.integer(edattain_string == "unknown"),
    married_cohab = as.integer(marst == 2 & !is.na(marst))
  )

# Standardize country names
intl_df <- intl_df %>%
  mutate(
    country_string = case_when(
      tolower(country_string) == "united states" ~ "US",
      tolower(country_string) == "puerto rico" ~ "PR",
      tolower(country_string) == "dominican republic" ~ "DR",
      tolower(country_string) == "el salvador" ~ "ES",
      tolower(country_string) %in% c("mexico", "honduras", "guatemala", "cuba", "colombia") ~
        str_to_title(country_string),
      TRUE ~ country_string
    ),
    year_str = as.character(year),
    country_year = paste0(country_string, "_", year_str),
    sex_string = tolower(as.character(as_factor(sex))),
    age2_string = as.character(as_factor(age2))
  )

# Calculate age weights for each country-year-sex combination
intl_age_weights <- intl_df %>%
  group_by(country_year, sex_string, age2, age2_string) %>%
  summarise(perwt_sum = sum(perwt, na.rm = TRUE), .groups = "drop") %>%
  group_by(country_year, sex_string) %>%
  mutate(
    total_perwt = sum(perwt_sum),
    age_weight_in = perwt_sum / total_perwt,
    match_var_in = paste0(country_year, sex_string, age2_string)
  ) %>%
  ungroup() %>%
  select(country_year, sex_string, age2, age2_string, age_weight_in, match_var_in)

# Save international weights
intl_age_weights %>%
  write_csv(paste0(weights_dir, "intl_age_weights.csv"))

# Load US age weights
us_weights_female <- read_csv(paste0(weights_dir, "us_age_weights_female.csv"), show_col_types = FALSE)
us_weights_male <- read_csv(paste0(weights_dir, "us_age_weights_male.csv"), show_col_types = FALSE)

us_weights <- bind_rows(
  us_weights_female %>% rename(age_weight_us = age_weight_female),
  us_weights_male %>% rename(age_weight_us = age_weight_male)
)

# Create match variables and merge weights
intl_df <- intl_df %>%
  mutate(
    match_var_in = paste0(country_year, sex_string, age2_string),
    match_var_us = paste0(sex_string, age2_string)
  )

# Merge US weights
intl_df <- intl_df %>%
  left_join(us_weights %>% select(match_var_us, age_weight_us), by = "match_var_us")

# Merge international weights
intl_df <- intl_df %>%
  left_join(intl_age_weights %>% select(match_var_in, age_weight_in), by = "match_var_in")

# Calculate age-standardized weights
intl_df <- intl_df %>%
  mutate(
    age_weight_ratio = age_weight_us / age_weight_in,
    perwt_age_standardized = perwt * age_weight_ratio
  )

# Keep analysis sample
intl_df <- intl_df %>%
  filter(country_year %in% c("Cuba_2012", "DR_2010", "Mexico_2010", "PR_2010", "Mexico_2020", "PR_2020"))

# Add secondary_or_above for consistency
intl_df <- intl_df %>%
  mutate(secondary_or_above = as.integer(secondary_completed == 1 | university_completed == 1))

saveRDS(intl_df, paste0(data_dir, "CuDrMePrUs_10_12.rds"))
write_dta(intl_df, paste0(data_dir, "CuDrMePrUs_10_12.dta"))
cat("Saved international comparison data:", nrow(intl_df), "observations\n")

cat("\nData preprocessing complete!\n")
cat("Output files:\n")
cat("  - US_2010_v100.rds / .dta\n")
cat("  - US_2020_v100.rds / .dta\n")
cat("  - US_2010_v100_census.rds / .dta\n")
cat("  - CuDrMePrUs_10_12.rds / .dta\n")
```

# Session Info

```{r session_info}
sessionInfo()
```
